---
title: 'PHASE 1: DATA GATHERING'
subtitle: "Thesis TUe - Automated Valuation Model for Commercial Real Estate"
author: "Bas Hilgers | Data: Cushman&Wakefield (confidential)"
date: "August 2018"
output:
  html_document:
    toc: yes
    toc_depth: 4
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<a id="top"></a>

---

# 1. INTRODUCTION

<br>

In the first phase of our Analytical Workflow we explain how we *gather data* from a variety of (semi) public sources. The codes below will make use of APIs, personal (Access) databases and GIS. 

*Note*: Some sources are invalid as is. Customize to functions found in **R_code_API_requests.Rmd**.

First we specify the directory into which the downloaded data will be stored, unzipped, etc. In addition we specifiy some necessary options for the process.

```{r eval=FALSE}

  ## libraries
  library(knitr)   # publication options markdown
  library(RODBC)   # database connection
  library(rgdal)   # GIS
  library(dplyr)   # data manipulation
  library(plyr)    # data manipulation

```

Set directories and global options.

```{r set-options}

  ## Set data and code directory
  wd.dir   <- file.path('C:/Users/bashi/Desktop/TUe Thesis')
  data.dir <- file.path(wd.dir, '5 Data', 'AVM database')
  code.dir <- file.path(wd.dir, '6 Modelling', 'modules')
  
  ## file options
  options(width=120) # control width output
  knitr::opts_chunk$set(out.width='800px', dpi=200) # control width figures
  options(scipen=999)
  options(stringsAsFactors = FALSE)
  
```

Next we load a set of **custom functions** that are used during the data gathering phase.

```{r eval=FALSE}

  source(paste0(code.dir, '/custom_functions_data_gathering.R'))

```

We load the raw **input data** into memory that contains the transactions including an list of addressses for which additional data needs to be gathered.

```{r eval=FALSE}

  ## establish db connection
  db.conn <- odbcConnectAccess2007(file.path(data.dir, 'AVM_database.accdb'))

  ## Read in sales data
  df_XL <- sqlFetch(db.conn, "tbl2_data_managed")
  df_XL$comments_transaction <- NULL # too large datafields

```

We divide the data into Chunks in order to limit our data as we are bound by 1000 requests per day due to our API key and preview our data.

```{r eval=FALSE}

  ## chunck raw data
  df_XL <- df_XL %>% slice(1:1000)
  df_XL <- df_XL %>% dplyr::arrange(BOG_ID)

  ## preview data structure
  names(df_XL)
                                      
```

We subset the data to include only necessary variables. Observations that are not useable (yet) are ecxluded from the data. That is, if they don't have an location or data. The missing data is saved as a .csv for further exploration.

```{r eval=FALSE}

  # select complete datapoints only
  df_XL <- dplyr::select(.data=df_XL, BOG_ID, housenumber, postcode)

  ## exclude missing data 
  df_XL <- df_XL %>% dplyr::filter(!is.na(BOG_ID) & !is.null(BOG_ID ) & BOG_ID != "")
  df_XL <- df_XL %>% dplyr::filter(!is.na(housenumber) & !is.null(housenumber) & housenumber != "")
  df_XL <- df_XL %>% dplyr::filter(!is.na(postcode) & !is.null(postcode) & postcode != "")
  
  head(df_XL)

```
 
<br>

---

# 2. DATA GATHERING

<br>

### A. BAG information

First, we obtain *geographical information* about the properties through the BAG API. Variables obtained are 'longitude' and 'lattitude' plus the links to 'openbare ruimte' and 'pandrelatering'.

```{r eval=FALSE, message=FALSE}  

  ## set source of GET fuction
  source(file.path(code.dir, 'GET_BAG_LL.R'))

  ## GET verblijfsobject data
  OUTPUT_verblijfsobject <- GET_verblijfsobject(df_XL$housenumber, df_XL$postcode)
  
  ## append data
  df_XL$longitude <- OUTPUT_verblijfsobject$lng
  df_XL$latitude <- OUTPUT_verblijfsobject$lat
  
```

Second, we get information about the *public space*. Variables obtained are the 'streetname' and 'city'.

```{r eval=FALSE}  

  ## set source of GET fuction
  source(file.path(code.dir, 'GET_BAG_public_space.R'))

  ## GET verblijfsobject data
  OUTPUT_public_space <- GET_public_space(OUTPUT_verblijfsobject$link_openbareruimte)
  
  ## append data
  df_XL$streetname <- OUTPUT_public_space$street
  df_XL$city <- OUTPUT_public_space$city
  
```

Rename 's Gravenhage to 'Den Haag'

```{r eval=FALSE} 

  df_XL$city <- gsub("'s-Gravenhage", "Den Haag", df_XL$city)

```

Third, we obtain the *building info*. Variables obtained are the 'BAG_object_ID' and 'construction_year'.

```{r eval=FALSE}  

  ## set source of GET fuction
  source(file.path(code.dir, 'GET_BAG_building_info.R'))

  ## GET verblijfsobject data
  OUTPUT_BAG_object <- GET_object_info(OUTPUT_verblijfsobject$link_pandrelatering)
  
  ## append data
  df_XL$BAG_object_ID <- OUTPUT_BAG_object$BAG_object_ID
  df_XL$BAG_construction_year <- OUTPUT_BAG_object$construction_year 
  
```


Last, we clear some of our memory in R and show data structure.

```{r eval=FALSE}  

  ## clean memory
  remove(OUTPUT_BAG_object)
  remove(OUTPUT_public_space)
  remove(OUTPUT_verblijfsobject)
  
  ## preview data
  str(df_XL)
  
```

<br>

### B. Region info

We obtain information about the regions in which the property is situated that is often used as an ID. The data is stored in an access database and is obtained based on the PC6 code. If not found, use broader pc4 code.

```{r eval=FALSE}  

  ## establish db connection
  db.region <- odbcConnectAccess2007(file.path(data.dir, 'Regions_data.accdb'))

  for(i in 1:nrow(df_XL)){
    
    ## query
    query <- sprintf("SELECT * FROM regions_NUTS WHERE postcode6 = '%s';", df_XL$postcode[i])
    data <- sqlQuery(db.region , query)
    
    ## append data
    if (dim(data)[1] != 0){
      df_XL$NUTS1_ID[i] <- as.character(data$NUTS1_ID)
      df_XL$NLD_region[i] <- as.character(data$Lands)
      df_XL$NUTS2_ID[i] <- as.character(data$NUTS2_ID)
      df_XL$COROP_region[i] <- as.character(data$COROP_region)
      df_XL$NUTS3_ID[i] <- as.character(data$NUTS3_ID)
      df_XL$provinciecode[i] <- as.character(data$provinciecode)
      df_XL$provincie[i] <- as.character(data$provincie)
      df_XL$gemeentecode[i] <- as.character(data$gemeentecode)
      df_XL$gemeente[i] <- as.character(data$gemeente)
      df_XL$wijkcode[i] <- as.character(data$wijkcode)
      df_XL$wijk[i] <- as.character(data$wijk)
      df_XL$buurtcode[i] <- as.character(data$buurtcode)
      df_XL$buurt[i] <- as.character(data$buurt)
    }
    
    ## clean memory
    remove(data)
    remove(query)
  }

  ## close connection
  odbcClose(db.region)
  remove(db.region)

```


<br>

### C. CW region

Next, load spatial polygon data.

```{r eval=FALSE}  

  ## load GIS (polygons and distict info)
  GIS_kantoren <- readOGR(dsn=file.path(data.dir,'CW.gdb'), "Kantoren", verbose=FALSE)
  GIS_bedrijfsruimte <- readOGR(dsn=file.path(data.dir,'CW.gdb'), "Bedrijfsruimte", verbose=FALSE)
  
  for(i in 1:nrow(df_XL)){
    if(!is.na(df_XL$lng[i]) && !is.na(df_XL$lat[i])){
    
      ## create spatial point for the subject property
      coords = cbind(df_XL$lng[i], df_XL$lat[i])
      sp = SpatialPoints(coords)
      
      ## allign projections
      proj4string(sp) <- proj4string(GIS_kantoren)
      
      ## check whether point subject property falls within polygon CW regions
      DLF_naam <- over(sp, GIS_kantoren) 
      if(!is.na(DLF_naam$FileName[1])) {
        
        ## append data kantoorgebied ## rename to CW_district..
        df_XL$CW_district_ID[i] <- as.character(DLF_naam$Deelgebied_id[1]) 
        df_XL$CW_district_name[i] <- as.character(DLF_naam$Deelgebied_naam[1])
        df_XL$CW_district_type[i] <- as.character(DLF_naam$Segment[1])     
      
      } else {
        
        # check bedrijventerreinen
        DLF_naam <- over(sp, GIS_bedrijfsruimte) 
        
        ## append data bedrijfsterrein
        df_XL$CW_district_ID[i] <- as.character(DLF_naam$DLG_Num[1])
        df_XL$CW_district_name[i] <- as.character(DLF_naam$DLG_Naam[1])
        df_XL$CW_district_type[i] <- as.character(DLF_naam$Segment[1]) 
        
      }

      ## clean memory
      remove(DLF_naam)
    }
  }
  
  ## replace NA with OTHER
  df_XL$CW_district_ID[is.na(df_XL$CW_region_ID)] <- "OTHER"
  df_XL$CW_district_name[is.na(df_XL$CW_region_name)] <- "OTHER"
  df_XL$CW_district_type[is.na(df_XL$CW_region_segment)] <- "OTHER"

  ## clean memory
  remove(sp)
  remove(GIS_bedrijfsruimte)
  remove(GIS_kantoren)
  remove(coords)
  
  ## what to do if both KA and BE? GE, but which name?

```

Last, we check wheter the city is large / small

```{r eval=FALSE}

  ## city categories
  city_cat <- c("Amsterdam", "Amsterdam-Zuidoost","Amsterdam_duivendrecht", "Schiphol-Rijk", 
                "Hoofddorp", "Rotterdam", "Den Haag", "Utrecht", "Eindhoven")
  df_XL <- df_XL %>% rowwise() %>%
           dplyr::mutate(city_category = ifelse(city %in% city_cat,"large", "small"))

```

<br>

### D. Energie prestatie score

Get EP data. If not found, search for closest building to number that has highest probability to show a similar label.

```{r eval=FALSE}  

  ## establish db connection
  db.ep <- odbcConnectAccess2007(file.path(data.dir, 'energie_prestatie.accdb'))

  for(i in 1:nrow(df_XL)){
    
    ## query
    query <- sprintf("SELECT TOP 1 pand_postcode, pand_huisnummer, pand_energieklasse, 
                                   pand_gebouwtype, pand_gebouwsubtype
                      FROM V20180501_csv
                      WHERE pand_postcode = '%s'
                      ORDER BY ABS( pand_huisnummer - %g );",
                      df_XL$postcode[i], df_XL$housenumber[i])

    ## get data
    data <- sqlQuery(db.ep , query)

    ## append data
    if (length(data)[1] != 2){
    df_XL$EP_huisnummer[i] <- as.numeric(data$pand_huisnummer[1])
    df_XL$EP_postcode[i] <- as.numeric(data$pand_postcode[1])
    df_XL$EP_energieklasse[i] <- as.character(data$pand_energieklasse[1])
    df_XL$EP_gebouwtype[i] <- as.character(data$pand_gebouwtype[1])
    df_XL$EP_gebouwsubtype <- as.character(data$pand_gebouwsubtype[1])
    }
    ## clean memory
    remove(data)
    remove(query)
  }
  
  ## data mutations
  df_XL$EP_energieklasse <- gsub(FALSE, "F", df_XL$EP_energieklasse)
  
  ## close connection
  odbcClose(db.ep)
  remove(db.ep)

```

<br>

### E. Building height

We obtain de building heigts from our personal database. Data obtained from *3D gebouwhoogte NL* from PDOK.

```{r eval=FALSE}  

  ## establish db connection
  db.height <- odbcConnectAccess2007(file.path(data.dir, 'building_heights.accdb'))

  for(i in 1:nrow(df_XL)){
    if(!is.na(df_XL$BAG_object_ID[i])){
      
      ## query
      query <- sprintf("SELECT * FROM building_heights_NLD WHERE ID = '%s';", 
                       as.numeric(df_XL$BAG_object_ID[i]))
      data <- sqlQuery(db.height , query)
      
      ## append data
      if(dim(data)[1] != 0){
        if(data$height_max[1] < data$height_avg[1]){
          df_XL$building_height[i] <- data$height_avg[1]
        } else {
          df_XL$building_height[i] <- data$height_max[1]
        }
      }
      ## clean memory
      remove(data)   
      remove(query)
      
    }
  }

  ## close connection
  odbcClose(db.height)
  remove(db.height)
  

```

<br>

### F. Leefbaarometer

We obtain de leefbaarometer from our personal database that is constructed with data from *CBS*.

```{r eval=FALSE}  

  ## establish db connection
  db.lbm <- odbcConnectAccess2007(file.path(data.dir, 'leefbaarometer.accdb'))

  for(i in 1:nrow(df_XL)){
    if(!is.na(df_XL$buurtcode[i])){
      
      ## select relevant year
      #year = substring(df_XL$date_transaction[i],7,10)
      year = as.numeric(substr(df_XL$transaction_date[i], 1, 4))
      if (is.na(year)) { year = df_XL$year[(i-1)] } # previous observation (observed around same period)
      if (is.na(year)) { year = 2016 } # no previous than 2016
      if (year > 2014) {year = 2016
      } else if (year > 2012) {year = 2014
      } else (year = 2012)
      
      ## query buurt
      query <- sprintf("SELECT * FROM buurt_2016_stand_panel WHERE GBD = '%s' AND JAAR = %d;", 
                       as.character(df_XL$buurtcode[i]), as.numeric(year))
      data <- sqlQuery(db.lbm , query)
      
        ## if no buurt score take try broader neighborhood score
        if(is.na(data$RLBRMTR[1])) {
          query <- sprintf("SELECT * FROM wijk_2016_stand_panel WHERE GBD = '%s' AND JAAR = %d;", 
                           as.character(df_XL$wijkcode[i]), as.numeric(year))
          data <- sqlQuery(db.lbm , query)        
        }
      
        ## if no neighborhood score take broader municipality score
        if(is.na(data$RLBRMTR[1])) {
          query <- sprintf("SELECT * FROM gemeente_2016_stand_panel WHERE GBD = '%s' AND JAAR = %d;", 
                           as.character(df_XL$gemeentecode[i]), as.numeric(year))
          data <- sqlQuery(db.lbm , query)        
        }
      
      ## append data
      df_XL$lbm_district[i] <- as.character(data$GBD)[1]
      df_XL$lbm_year[i] <- as.numeric(data$JAAR)[1]
      df_XL$lbm_total_score[i] <- as.numeric(data$RLBRMTR)[1]
      df_XL$lbm_woningen[i] <- as.numeric(data$RLBWON)[1]
      df_XL$lbm_bewoners[i] <- as.numeric(data$RLBBEV)[1]
      df_XL$lbm_voorzieningen[i] <- as.numeric(data$RLBVRZ)[1]
      df_XL$lbm_veiligheid[i] <- as.numeric(data$RLBVEI)[1]
      df_XL$lbm_fysieke_omgeving[i] <- as.numeric(data$RLBFYS)[1]
      
      ## clean memory
      remove(data)
      remove(query)  
      remove(year)
      
    }
  }

  ## close connection
  odbcClose(db.lbm)
  remove(db.lbm)


```

<br>

### G. Walkscore

Next, get walkscore with use of the API.

```{r eval=FALSE}  

  ## Source of function
  source(file.path(code.dir, 'GET_walkscore.R'))

  ## GET data
  OUTPUT_walkscore <- GET_walkscore(address_number = df_XL$address_number, 
                                    streetname = df_XL$streetname, 
                                    postcode = df_XL$postcode,
                                    city = df_XL$city, 
                                    lat = df_XL$lat,
                                    lng = df_XL$lng)

  ## append data
  df_XL$walkscore <- OUTPUT_walkscore$walkscore
  df_XL$walkscore_date <- OUTPUT_walkscore$walkscore_date
  df_XL$walkscore_link <- OUTPUT_walkscore$walkscore_link
  
  ## clean memory
  remove(OUTPUT_walkscore)
  
```

<br>

### H. Nearest highway access

First, obtain de (geo)location of the 2 nearest highway exits based on distance (pythagoras). 
Note: the closest in distance doesn't have to be the fastest!

```{r eval=FALSE}  

  ## establish db connection
  db.highway <- odbcConnectAccess2007(file.path(data.dir, 'transportation.accdb'))

  for(i in 1:nrow(df_XL)){
    if(!is.na(df_XL$lat[i]) && !is.na(df_XL$lng[i])){
      
      ## query
      query <- sprintf("SELECT TOP 2 name, lat, lng
                        FROM highway_exits
                        ORDER BY sqr(((lat-%s)^2+(lng-%s)^2));", 
                        df_XL$lat[i], df_XL$lng[i])
      
      ## get query data
      data <- sqlQuery(db.highway , query)
      
      ## append data
      df_XL$highway_name1[i] <- as.character(data$name[1])
      df_XL$highway_lat1[i] <- data$lat[1]
      df_XL$highway_lng1[i] <- data$lng[1]
      df_XL$highway_name2[i] <- as.character(data$name[2])
      df_XL$highway_lat2[i] <- data$lat[2]
      df_XL$highway_lng2[i] <- data$lng[2]
      
      ## clean memory
      remove(data)
      remove(query)
      
    }
  }

  ## close connection
  odbcClose(db.highway)
  remove(db.highway)
  
```

Second, we obtain info about the duration and distance and select the option that is closest timewise:

```{r eval=FALSE}  

  ## set source of GET fuction
  source(file.path(code.dir, 'GET_google_distances_duration.R'))

  ## GET verblijfsobject data
  OUTPUT_google_duration <- GET_google_distances_duration(lat_object = df_XL$lat,
                                                          lng_object = df_XL$lng,
                                                          lat_place = df_XL$highway_lat1,
                                                          lng_place = df_XL$highway_lng1,
                                                          mode = "driving")
  
  ## append data
  df_XL$highway_distance1 <- OUTPUT_google_duration$distance
  df_XL$highway_duration1 <- OUTPUT_google_duration$duration
  
  ## GET verblijfsobject data
  OUTPUT_google_duration <- GET_google_distances_duration(lat_object = df_XL$lat,
                                                          lng_object = df_XL$lng,
                                                          lat_place = df_XL$highway_lat2,
                                                          lng_place = df_XL$highway_lng2,
                                                          mode = "driving")
  
  ## append data
  df_XL$highway_distance2 <- OUTPUT_google_duration$distance
  df_XL$highway_duration2 <- OUTPUT_google_duration$duration
  
  
  ## nearest highway among 2 options
    for(i in 1:nrow(df_XL)){
      if(!is.na(df_XL$highway_duration2[i])){
      if(df_XL$highway_duration1[i] < df_XL$highway_duration2[i]){
        df_XL$highway_access_name[i] <- df_XL$highway_name1[i]
        df_XL$highway_access_lat[i] <- df_XL$highway_lat1[i]
        df_XL$highway_access_lng[i] <- df_XL$highway_lng1[i]
        df_XL$highway_access_distance[i] <- df_XL$highway_distance1[i]
        df_XL$highway_access_duration[i] <- df_XL$highway_duration1[i]
      } else {
        df_XL$highway_access_name[i] <- df_XL$highway_name2[i]
        df_XL$highway_access_lat[i] <- df_XL$highway_lat2[i]
        df_XL$highway_access_lng[i] <- df_XL$highway_lng2[i]
        df_XL$highway_access_distance[i] <- df_XL$highway_distance2[i]
        df_XL$highway_access_duration[i] <- df_XL$highway_duration2[i]
      }
      }
    }
  
  ## remove temps
  df_XL$highway_name1 <- NULL
  df_XL$highway_lat1 <- NULL
  df_XL$highway_lng1 <- NULL
  df_XL$highway_distance1 <- NULL
  df_XL$highway_duration1 <- NULL
  df_XL$highway_name2 <- NULL
  df_XL$highway_lat2 <- NULL
  df_XL$highway_lng2 <- NULL
  df_XL$highway_distance2 <- NULL
  df_XL$highway_duration2 <- NULL
  
  ## clean memory
  remove(OUTPUT_google_duration)
  
```

<br>

### I. Nearest station

First, obtain de (geo)location of the nearest train station through google API.

```{r eval=FALSE}  

  ## set source of GET fuction
  source(file.path(code.dir, 'GET_google_places.R'))

  ## GET verblijfsobject data
  OUTPUT_nearest_station <- GET_google_places(lat = df_XL$lat,
                                              lng = df_XL$lng,
                                              places_type = "train_station")
  
  ## correct names
  OUTPUT_nearest_station$name <- gsub(" railway station", "", OUTPUT_nearest_station$name)
  
  ## append data
  df_XL$train_station_name <- OUTPUT_nearest_station$name
  df_XL$train_station_lat <- OUTPUT_nearest_station$X_lat
  df_XL$train_station_lng <- OUTPUT_nearest_station$X_lng
  
  ## clean memory
  remove(OUTPUT_nearest_station)
  
  ### TO DO: waarom DH centraal rare naam?
  
```

Second, we obtain additional data from our personal database. Mainly wheter the station is an *intercity station*.

```{r eval=FALSE}  

  ## establish db connection
  db.trans <- odbcConnectAccess2007(file.path(wd.dir, "data", 'transportation.accdb'))

  for(i in 1:nrow(df_XL)){
    if(!is.na(df_XL$train_station_name[i])){
    
      ## query
      query <- sprintf("SELECT * FROM train_stations WHERE train_station_name = '%s';", 
                       as.character(df_XL$train_station_name[i]))
      data <- sqlQuery(db.trans , query)
      
      ## append data
      if(is.null(dim(data))) {
      }else{
      if(dim(data)[1] != 0) {
        df_XL$train_station_intercity[i] <- TRUE
      } else {
        df_XL$train_station_intercity[i] <- FALSE
      }
      }
      
      ## clean memory
      remove(query)
      remove(data)
    }
  }

  ## close connection
  odbcClose(db.trans)
  remove(db.trans)
  
```

Last, we obtain info about the duration and distance:

```{r eval=FALSE}  

  ## set source of GET fuction
  source(file.path(code.dir, 'GET_google_distances_duration.R'))

  ## GET verblijfsobject data
  OUTPUT_google_duration <- GET_google_distances_duration(lat_object = df_XL$lat,
                                                          lng_object = df_XL$lng,
                                                          lat_place = df_XL$station_lat,
                                                          lng_place = df_XL$station_lng,
                                                          mode = "walking")
  
  ## append data
  df_XL$train_station_distance <- OUTPUT_google_duration$distance
  df_XL$train_station_duration <- OUTPUT_google_duration$duration
  
  ## clear memory
  remove(OUTPUT_google_duration)
  
```

Last, we get the centrality indicator

```{r eval=FALSE}  

  ## centrality
  df_XL <- df_XL %>% rowwise() %>%
           dplyr::mutate(centrality = ifelse(train_station_distance <= 400 & 
                                             train_station_intercity %in% "TRUE", "central", "decentral"))

```

<br>

### J. Other OV

Other public transportation including bus, metro and tram.

```{r eval=FALSE} 

    ## Subway Station
    source(file.path(code.dir, 'GET_google_places.R'))
    OUTPUT_nearest_station <- GET_google_places(lat = df_XL$latitude,
                                                lng = df_XL$longitude,
                                                places_type = "subway_station")
      df_XL$subway_station_name <- OUTPUT_nearest_station$name
      df_XL$subway_station_lat <- OUTPUT_nearest_station$X_lat
      df_XL$subway_station_lng <- OUTPUT_nearest_station$X_lng
      
    source(file.path(code.dir, 'GET_google_distances_duration.R'))
    OUTPUT_google_duration <- GET_google_distances_duration(lat_object = df_XL$latitude,
                                                            lng_object = df_XL$longitude,
                                                            lat_place = df_XL$subway_lat,
                                                            lng_place = df_XL$subway_lng,
                                                            mode = "walking")
      df_XL$subway_station_distance <- OUTPUT_google_duration$distance
      df_XL$subway_station_duration <- OUTPUT_google_duration$duration


```

```{r eval=FALSE} 

    ## Bus Station
    source(file.path(code.dir, 'GET_google_places.R'))
    OUTPUT_nearest_station <- GET_google_places(lat = df_XL$latitude,
                                                lng = df_XL$longitude,
                                                places_type = "bus_station")
      df_XL$bus_station_name <- OUTPUT_nearest_station$name
      df_XL$bus_station_lat <- OUTPUT_nearest_station$X_lat
      df_XL$bus_station_lng <- OUTPUT_nearest_station$X_lng
      
    source(file.path(code.dir, 'GET_google_distances_duration.R'))
    OUTPUT_google_duration <- GET_google_distances_duration(lat_object = df_XL$latitude,
                                                            lng_object = df_XL$longitude,
                                                            lat_place = df_XL$bus_station_lat,
                                                            lng_place = df_XL$bus_station_lng,
                                                            mode = "walking")
      df_XL$bus_station_distance <- OUTPUT_google_duration$distance
      df_XL$bus_station_duration <- OUTPUT_google_duration$duration

```

```{r eval=FALSE} 

    ## Airport
    source(file.path(code.dir, 'GET_google_places.R'))
    OUTPUT_nearest_station <- GET_google_places(lat = df_XL$latitude,
                                                lng = df_XL$longitude,
                                                places_type = "airport")
      df_XL$airport_name <- OUTPUT_nearest_station$name
      df_XL$airport_lat <- OUTPUT_nearest_station$X_lat
      df_XL$airport_lng <- OUTPUT_nearest_station$X_lng
      
    source(file.path(code.dir, 'GET_google_distances_duration.R'))
    OUTPUT_google_duration <- GET_google_distances_duration(lat_object = df_XL$latitude,
                                                            lng_object = df_XL$longitude,
                                                            lat_place = df_XL$airport_lat,
                                                            lng_place = df_XL$airport_lng,
                                                            mode = "drivint")
      df_XL$airport_distance <- OUTPUT_google_duration$distance
      df_XL$airport_duration <- OUTPUT_google_duration$duration

```

```{r eval=FALSE} 

    ## Transit station
    source(file.path(code.dir, 'GET_google_places.R'))
    OUTPUT_nearest_station <- GET_google_places(lat = df_XL$latitude,
                                                lng = df_XL$longitude,
                                                places_type = "transit_station")
      df_XL$transit_station_name <- OUTPUT_nearest_station$name
      df_XL$transit_station_lat <- OUTPUT_nearest_station$X_lat
      df_XL$transit_station_lng <- OUTPUT_nearest_station$X_lng
      
    source(file.path(code.dir, 'GET_google_distances_duration.R'))
    OUTPUT_google_duration <- GET_google_distances_duration(lat_object = df_XL$latitude,
                                                            lng_object = df_XL$longitude,
                                                            lat_place = df_XL$transit_station_lat,
                                                            lng_place = df_XL$transit_station_lng,
                                                            mode = "walking")
      df_XL$transit_station_distance <- OUTPUT_google_duration$distance
      df_XL$transit_station_duration <- OUTPUT_google_duration$duration
      

```

<br>

---

# 3. STORE DATA

<br>

And save to database

```{r eval=FALSE}

  try(sqlDrop(db.conn, "tbl1_data_gathered", errors = TRUE))
  sqlSave(db.conn, df_XL, tablename = "tbl1_data_gathered", append = FALSE,
          rownames = FALSE, colnames = FALSE, nastring = "") # cannot use NA in Access for numeric fields

```

Lastly, we close the database connection.

```{r eval=FALSE}

  odbcClose(db.conn)
  odbcCloseAll()
  remove(db.conn)
  
  # clean workspace
  rm(list = ls())

```

<br>

---

> Go back to [top](#top)